### 2019-10-24 Manuscript/manuscript.tex
add text referring to THO ecologies

There is strong evidence both in Laurentia's geological and paleomagnetic record for differential plate tectonic motion between 2.2 and 1.8 Ga. The continued history of accretionary orogenesis and the evaluation of Laurentia's pole path in comparison to other continents from 1.8 Ga onward supports the continual operation of plate tectonics throughout the rest of the Proterozoic and Phanerozoic as well. While this evidence fits with the majority of interpretations of the timing of initiation of modern-style plate tectonics (see summary in \citealp{Korenaga2013a}), there continue to be arguments proposing that a stagnant lid persisted through the Mesoproterozoic Era (1.6 to 1.0 Ga) and into the Neoproterozoic  with plate tectonics not initiating until ca. 0.8 Ga \citep{Hamilton2011a, Stern2018a}. These arguments rest largely on the relative lack of the types of low-temperature high-pressure metamorphic rocks such as blueschists that form in subduction zones in the Proterozoic relative to the Phanerozoic \citep{Stern2013a}. An alternative interpretation for this lack of blueschists in the Proterozoic is that such a shift in metamorphic regime is the predicted result of secular evolution of mantle chemistry rather than a harbinger of the onset of plate tectonics \citep{Palin2015a}. While this line of evidence is intriguing, to argue that there was not differential plate tectonic motion in the Paleoproterozoic and Mesoproterozoic is to ignore a vast breadth and depth of geological and paleomagnetic data. As an example, consider the record of Laurentia in the late Mesoproterozoic. As can be seen in Table 2, there are many high-quality paleomagnetic poles between 1110 and 1070 Ma. These poles constrain rapid motion of Laurentia leading up to collisional orogenesis as illustrated in Figure \ref{fig:Grenville_reconstructions} providing strong evidence for differential plate motion at the time.

>>>>

There is strong evidence both in Laurentia's geological and paleomagnetic record for differential plate tectonic motion between 2.2 and 1.8 Ga. The continued history of accretionary orogenesis and the evaluation of Laurentia's pole path in comparison to other continents from 1.8 Ga onward supports the continual operation of plate tectonics throughout the rest of the Proterozoic and Phanerozoic as well. While this evidence fits with the majority of interpretations of the timing of initiation of modern-style plate tectonics (see summary in \citealp{Korenaga2013a}), there continue to be arguments proposing that a stagnant lid persisted through the Mesoproterozoic Era (1.6 to 1.0 Ga) and into the Neoproterozoic  with plate tectonics not initiating until ca. 0.8 Ga \citep{Hamilton2011a, Stern2018a}. These arguments rest largely on the relative lack of the types of low-temperature high-pressure metamorphic rocks such as blueschists that form in subduction zones in the Proterozoic relative to the Phanerozoic \citep{Stern2013a}. An alternative interpretation for this lack of blueschists in the Proterozoic is that such a shift in metamorphic regime is the predicted result of secular evolution of mantle chemistry rather than a harbinger of the onset of plate tectonics \citep{Palin2015a}. While this line of evidence is intriguing, to argue that there was not differential plate tectonic motion in the Paleoproterozoic and Mesoproterozoic is to ignore a vast breadth and depth of geological and paleomagnetic data. From a paleomagnetic perspective there is strong paleomagnetic support for independent and differential motion of the Slave and Superior provinces as is illustrated in Figure \ref{fig:Superior_Slave_recons}. From a geological perspective, the the Trans-Hudson orogenic cycle, the Grenville orogenic cycle, and the Appalachian orogenic cycle are all well-explained with a mobilistic interpretation that includes phases of accretionary followed by collisional orogenesis (Fig. \ref{fig:tectonic_history}). One could counter that this perspective results from a plate-tectonic-centric viewpoint that lacks creativity to see the record as resulting from other processes than modern-style plate tectonics. However, in addition to the broad geological record showing an amalgamation of terranes as would be expected to arise through plate tectonics, there are also ecologites preserved in the Trans-Hudson orogen that preserve evidence for high-pressure/low-temperature metamorphic conditions ca. 1.8 Ga \citep{Weller2017a}. Similar to the Himalayan orogen, these rocks are interpreted to be the result of deep continental subduction and exhumation associated with convergent plate tectonics \citep{Weller2017a}. Outside of Laurentia, there are examples of ecologites with geochemical affinity to oceanic crust such as that documented in the ca. 1.9 Ga Ubendian Belt of the Congo craton \citep{Boniface2012a}.

As an additional example, consider the record of Laurentia in the late Mesoproterozoic. As can be seen in Table 2, there are many high-quality paleomagnetic poles between 1110 and 1070 Ma. These poles constrain rapid motion of Laurentia leading up to collisional orogenesis as illustrated in Figure \ref{fig:Grenville_reconstructions} providing strong evidence for differential plate motion at the time.
